Utilizing Parallel Workers: \\LLVM's Vectorization Plan

### 2024-11-22 paper/paper.tex
paper

The goal was to unite any auto-vectorization systems (SLP, Inner-/Outer-Loop) into a single flexible
system~\cite{llvmintrvplan,llvmvplanstate}, as seen in Figure~\ref{fig:vplan-future}. This system would use VPlans to represent different 
vectorization approaches, abstracted away from the underlying IR. Each VPlan would be able to 
calculate its cost and transform itself to modify the final IR. At first the Legality 
phase (yellow) would check if the scalar code would be legal to vectorize at all. If so it would 
create different initial VPlans, representing the different initial ways to vectorize. In the 
following Planning phase (magenta) these initial VPlans would be optimized using various 
VPlan-to-VPlan transformations. Those transformations would be supervised by the VPlan's cost model.
Finally, the best VPlan would be chosen and materialized/executed to modify the underlying IR.
The best VPlan might also be to not vectorize at all (VF $=1$).

@@

@@

\subsubsection{VPlan and Integration with LV}
The main effort of setting up the VPlan structure and integrating it first into LV has seen
tremendous effort over the last years.

\subsubsection{Outer Loop Vectorization}
Development around outer loop vectorization has slowed down a bit as focus shifted more on VPlan
internals and integration into LV.

>>>>

The goal was to unite any auto-vectorization systems (SLP, Inner-/Outer-Loop) into a single flexible
system~\cite{llvmintrvplan,llvmvplanstate}, as seen in Figure~\ref{fig:vplan-future}. This system would use VPlans to represent different 
vectorization approaches, abstracted away from the underlying IR. Each VPlan would be able to 
calculate its cost and transform itself to modify the final IR. At first the Legality 
phase (yellow) would check if the scalar code would be legal to vectorize at all and if so initial 
VPlans would be constructed. In the following Planning phase (magenta) these initial VPlans would 
be optimized using various VPlan-to-VPlan transformations. Those transformations would be supervised 
by the VPlan's cost model. Finally, the best VPlan would be chosen and materialized/executed, 
modifying the underlying IR. The best VPlan might also be to not vectorize at all (VF $=1$).

@@

In its current state (LLVM 19.1.4) VPlan is deeply integrated into LV.

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{images/inner-loop-vplan-hcfg.png}
  \caption{Hierarchical CFG of a VPlan modelling the vectorized version of the loop in 
  Figure~\ref{fig:inner-loop-vec}.}
  \label{fig:inner-loop-vplan-hcfg}
\end{figure*}

@@

\subsubsection{General VPlan and Integration with LV}
The main effort of setting up the VPlan structure and integrating it first into LV has seen
tremendous effort over the last years. Florian Hahn presented future directions for 
VPlan~\cite{llvmvplanupdate}, which mostly evolve around transforming more LV-based decisions into
explicit VPlan-to-VPlan transformations. Additionally, a lot of cost model related decisions 
still end up being influenced by the old legacy LV cost model, however, transition to a VPlan-based 
cost model has already started.

\subsubsection{Outer Loop Vectorization}
Development around outer loop vectorization has slowed down a bit as focus shifted more onto VPlan
internals and integration into LV. The inner loop path and outer loop path are currently still
seperated by the experimental \texttt{-enable-vplan-native-path} flag. Further developing the 
outer loop path and eventually merging it with the innermost loop path is planned.
